\begin{thebibliography}{100}

\bibitem{Bellman1997} Richard Bellman, \emph{Introduction to matrix analysis},
SIAM, Society for Industrial and Applied Mathematics; $2^{nd}$ edition (1997).

\bibitem{Bevington} Philip R. Bevington, \emph{Data Reduction and Error Analysis in the Physical Sciences},
McGraw-Hill (1969).

\bibitem{Chan2007} Raymond H. Chan, and Chen Greif, and Diane P. O'Leary,
\emph{Milestones in matrix computation: Selected works of Gene H. Golub, with commentaries}, Oxford University Press (2007).

\bibitem{Demmel1997}  James W. Demmel, 
\emph{Applied numerical linear algebra},
SIAM, Society for Industrial and Applied Mathematics (1997).

\bibitem{Golub1996} Gene H. Golub, and Charles Van Loan,
\emph{Matrix Computations}, $3^{rd}$ Edition.
Johns Hopkins University Press (1996).

\bibitem{Higham2008} Nicholas J. Higham,
\emph{Functions of Matrices: Theory and Computation}, SIAM, Society for Industrial and Applied Mathematics (2008).

\bibitem{Horn1990}  Roger A. Horn, and Charles R. Johnson, 
\emph{Matrix analysis},
Cambridge University Press (1990).

\bibitem{Horn1991} Roger A. Horn, and Charles R. Johnson,
\emph{Topics in Matrix analysis}, $3^{rd}$ Edition.
Cambridge University Press (1991).

\bibitem{Mercer} Idris C. Mercer
\emph{Finding “nonobvious” nilpotent matrices}, (2005) \\
\texttt{http://www.idmercer.com/nilpotent.pdf}

\bibitem{Laub2005} Alan J. Laub,
\emph{Matrix analysis for scientists and engineers}, SIAM, Society for Industrial and Applied Mathematics (2005).

\bibitem{Meyer2000} Carl D. Meyer, 
\emph{Matrix analysis and applied linear algebra},
SIAM, Society for Industrial and Applied Mathematics (2000).

\bibitem{Strang2005} Gilbert Strang,
\emph{Linear Algebra and Its Applications}, SIAM, Society for Industrial and Applied Mathematics (2005).

\bibitem{Trefethen2000} Lloyd N. Trefethen, and David Bau,
\emph{Numerical linear algebra}, SIAM, Society for Industrial and Applied Mathematics (2000).

\bibitem{Eric} Eric W. Weisstein, "Characteristic Polynomial", from MathWorld--A Wolfram Web Resource.\\ \texttt{http://mathworld.wolfram.com/CharacteristicPolynomial.html}

\end{thebibliography}
