%%%
\clearpage
\begin{thebibliography}{99}
%
\bibitem{Bevington} Bevington, Philip R.  \emph{Data Reduction and Error Analysis for the Physical Sciences, 1e}, McGraw-Hill, 1969.
%
\bibitem{Businger} Businger, P.A. \& Golub, G.H., Paul \emph{Algorithm 358: Singular value decomposition of a complex matrix}, Comm. Assoc. Comp. Mach., 12, pp. 564--565, 1969.%
%
\bibitem{DeMoor1} De Moor, Bart \& Zha, Hongyuan \emph{A Tree of Generalizations of the Ordinary Singular Value Decomposition}, Linear Algebra and its Applications \texbf{147}(1991),  469--500.
%
\bibitem{DeMoor2} De Moor, Bart \& Van Dooren, Paul \emph{Generalizations of the singular values and QR decompositions}, SIAM J. Matrix Anal. Appl., \texbf{13}:4, pp. 993-1014, 1992.
%
\bibitem{Golub1} Golub, Gene H. \& Kahan, W.  \emph{Calculating the singular values and pseudo-inverse of a matrix}, J. SIAM Numer. Anal., \texbf{2}(2), pp. 205-224, 1965.
%
\bibitem{Golub2} Golub, Gene H. \& Reinsch, C.  \emph{Singular value decomposition and least squares solutions}, Numer. Math.  \texbf{14}: 403--420, 1970.
%
\bibitem{Golub} Golub, Gene H. \& Van Loan, Charles F. \emph{Matrix Computations, 3e}, Johns Hopkins Press, 1996.
%
\bibitem{HJ1}Horn, Roger A. \& Johnson, Charles R. \emph{Matrix Analysis}, Cambridge University Press, 1990.
%
\bibitem{HJ2}Horn, Roger A. \& Johnson, Charles R. \emph{Topics in Matrix Analysis}, Cambridge University Press, 1994.
%
\bibitem{Laub} Laub, Alan J. \emph{Matrix Analysis for Scientists and Engineers}, SIAM, 2005.
%
\bibitem{Meyer} Meyer, Carl D. \emph{Matrix Analysis and Applied Linear Algebra}, SIAM, 2001.
%
\bibitem{Penrose} Penrose, Roger \emph{A generalized inverse for matrices} Proceedings of the Cambridge Philosophical Society \textbf{51}: 406лл413, 1955.

\bibitem{Knuth92} D.E. Knuth, \emph{Two notes on notation}, Amer.
Math. Monthly \textbf{99} (1992), 403--422.

\end{thebibliography} 


\endinput