\section{Unitary matrices}

Unitary matrices are easy to work with.
two views: a set of basis vectors, a process
unitary, orthogonal
\begin{equation}
  \U{}\U{*} = \I{}
\end{equation}
or
\begin{equation}
  \U{-1} = \U{*}
\end{equation}
Idempotent, index $k=2$.

Perhaps too much detail

spheres into spheres

%%
\subsection[Products of unitary matrices]{Products of unitary matrices are unitary}

An important realization is that the product of unitary matrices is also unitary. We will see that by looking at domain matrices as products of elementary unitary forms the \svdp \ looks much simpler.

Consider two conformable unitary matrices $\U{}\in\cmplx{m\times m}$ and $\V{}\in\cmplx{m\times m}$. Is there product unitary also?

Define
\begin{equation}
  \W{} = \U{}\V{*}.
\end{equation}
The question we are asking is this: Does the matrix product
\begin{equation}
  \W{}\W{}^{*} = \W{}^{*}\W{} = \I{m}?
  \label{eq:domains:unitarytest}
\end{equation}

To be explicit we show
\begin{equation}
  \W{}^{*}= \paren{\U{}\V{*}}^{*} = \paren{\V{*}}^{*}\U{*} = \V{}\U{*}.
\end{equation}
This implies the following:
\begin{equation}
  \begin{split}
    \W{}\W{}^{*} &= \paren{\U{}\V{*}}\paren{\V{}\U{*}}\\
      &=\U{} \paren{\V{*}\V{}} \U{*}\\
      &= \U{}\,\U{*}\\
      &= \I{m}.
  \end{split}
\end{equation}
A similar process shows that
\begin{equation}
  \W{}^{*}\W{} = \I{m}.
\end{equation}
Therefore the criteria of equation \eqref{eq:domains:unitarytest} are satisfied. Therefore the product of two unitary matrices is unitary. Therefore the product of three or more matrices are unitary.

This shows that a unitary matrix may be the composition of other unitary matrices. The implication is that we may examine the domain matrices and reveal a composition of elementary processes.

\endinput