\section{Is the SVD unique?}
Is the SVD unique? The matrix of singular values $\sig{}$ is always unique. This implies that the embedded diagonal submatrix $\SS{}$ is unique. But the domain decomposition matrices $\X{}$ and $\Y{}$ are, in general, not unique.

\subsection{The orthogonal complement}
Think in the familiar world of $\real{3}$ and consider the perpendicular component of the $z-$axis, the $x-y$ plane.

The most natural decomposition is along the unit vectors $\hat{x}$ and $\hat{y}$. Of course and set of perpendicular vectors in the plane will work. For example, these vectors could be rotated by an arbitrary angle $\theta$ like so:
\begin{equation}
  \R{}(\theta)\mat{c|c}{\hat{x}&\hat{y}}=
  \mat{rr}
  {
   \cos \theta & -\sin \theta \\ \sin \theta & \cos \theta
   }
   \itwo =
  \mat{rr}
  {
   \cos \theta & -\sin \theta \\ \sin \theta & \cos \theta
   } =
   \mat{c|c}{\hat{x}'&\hat{y}'}.
\end{equation}
Reflections of the coordinates are also suitable decompositions. We could reflect through the $y-$axis
\begin{equation}
  \mat{c|c}{\hat{x}&\hat{y}} =
  \mat{rr}{-1&0\\0&1}\mat{c|c}{\hat{x}&\hat{y}} = \mat{c|c}{-\hat{x}'&\hat{y}'},
\end{equation}
reflect through the $x-$axis
\begin{equation}
  \mat{c|c}{\hat{x}&\hat{y}} =
  \mat{rr}{1&0\\0&-1}\mat{c|c}{\hat{x}&\hat{y}} = \mat{c|c}{\hat{x}'&-\hat{y}'},
\end{equation}
reflect through the line $y=x$
\begin{equation}
  \mat{c|c}{\hat{x}&\hat{y}} =
  \mat{rr}{-1&0\\0&-1}\itwo = \mat{c|c}{-\hat{x}&-\hat{y}},
\end{equation}
or reflect through a line of arbitrary slope through the origin.

But these are all operations on the orthogonal complement which are stenciled out by the $\sig{}$ matrix.

What about the active components? Are there transformation matrices $\T{y}$ and $\T{x}$ such that
\begin{equation}
  \svdax{*} = \T{y}\,\Y{}\,\sig{}\,\T{x}\,\X{}?
\end{equation}

Isn't this equivalent to
\begin{equation}
  \T{y}\,\Y{}\,\sig{}\,\T{x}=\I{n}?
\end{equation}



\endinput