\documentclass[10pt]{newsiambook}

\usepackage{amsmath,amssymb}
\usepackage{color}
\usepackage{colortbl}
\usepackage{graphicx}
%\usepackage{pifont}
\usepackage{makeidx}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{crop}
\usepackage{pdfpages}
\usepackage{pdflscape}
\usepackage{tabularx}
\usepackage{pstricks}
\setlength\minrowclearance{4pt}
\usepackage{sidecap}
%\usepackage{cancel}
 
\crop	
\newcommand{\pathname}{common/}
\newcommand{\fullpath}{\pathname}

\input{\pathname declarations.tex}
	
\makeindex

\begin{document}

\frontmatter
%
\tableofcontents
%
\listoffigures
%%
\listoftables
%
%\input{chapters/frontmatter/preface}
%\input{chapters/frontmatter/literature}

\mainmatter
%%%%%%%
\part{Introducing the SVD}
\input{chapters/ftola/"head ftola"}
\input{chapters/"svd ex nihilo"/"head svd ex nihilo"}
\input{chapters/"computing the svd I"/"head computing the svd I"}
\input{chapters/"computing the svd II"/"head computing the svd II"}

%%%%%%%
\part{A closer look at the SVD}
\input{chapters/"computing the svd III"/"head computing the svd III"}
\input{chapters/proofs/"head proofs"}
%
%%%%%%%%
\part{Applications}
\input{chapters/"least squares"/"head least squares"}
\input{chapters/"least squares"/"head least squares II"}
\input{chapters/"pseudoinverse"/"head pseudoinverse"}
\input{chapters/projectors/"head projectors"}
%
%%%%%%%%%
\part{Seeing the SVD}
\input{chapters/bases/"head bases"}
%\input{chapters/"information content I"/"head information content I"}
%\input{chapters/"information content Ia"/"head information content Ia"}
%\input{chapters/"information content Ib"/"head information content Ib"}
%\input{chapters/"information content Ic"/"head information content Ic"}
%\input{chapters/"information content II"/"head information content II"}
%\input{chapters/"information content III"/"head information content III"}
%
%%%%%%%%%
\part{Explorations}
%%%\input{chapters/diagonalization/"head diagonalization"}
%\input{chapters/pantheon/"head pantheon"}
\input{chapters/examples/"head examples"}
%%%\input{chapters/topics/"head topics"}
%
%%%%%%%%%
\part{Machinations}
%%\input{chapters/proofs/"head proofs"}
%\input{chapters/"matrix analysis"/"head matrix analysis"}
%
%
%%%%%%%%%
%\part{Peroration}
%%\input{chapters/proofs/"head proofs"}
%
%%%%%%%%
\part{Appendices}
\appendix
%%\input{chapters/appendices/summaries/"head summaries"}
\input{chapters/appendices/history/"head history"}
\input{chapters/appendices/precis/"head precis"}
\input{chapters/appendices/"building blocks"/"head building blocks"}
\input{chapters/appendices/"decomposition summaries"/"head decomposition summaries"}
\input{chapters/appendices/"gram schmidt"/"head gram schmidt"}
\input{chapters/appendices/jargon/"head jargon"}
\input{chapters/appendices/complex/"head complex"}
\input{chapters/appendices/definitions/"head definitions"}

%%%%%%
\part{\ \ Backmatter}
\backmatter
%\input{backmatter/acronyms}
\input{chapters/backmatter/notation/"head notation"}
%%\input{backmatter/glossary}
%\bibliography{chapters/backmatter/bibliography/svdfp}
%\input{chapters/backmatter/bibliography/svdfp.bib}

\begin{thebibliography}{100}

\bibitem{Bellman1997} Richard Bellman, \emph{Introduction to matrix analysis},
SIAM, Society for Industrial and Applied Mathematics; $2^{nd}$ edition (1997).

\bibitem{Bevington} Philip R. Bevington, \emph{Data Reduction and Error Analysis in the Physical Sciences},
McGraw-Hill (1969).

\bibitem{Chan2007} Raymond H. Chan, and Chen Greif, and Diane P. O'Leary,
\emph{Milestones in matrix computation: Selected works of Gene H. Golub, with commentaries}, Oxford University Press (2007).

\bibitem{Demmel1997}  James W. Demmel, 
\emph{Applied numerical linear algebra},
SIAM, Society for Industrial and Applied Mathematics (1997).

\bibitem{Golub1996} Gene H. Golub, and Charles Van Loan,
\emph{Matrix Computations}, $3^{rd}$ Edition.
Johns Hopkins University Press (1996).

\bibitem{Higham2008} Nicholas J. Higham,
\emph{Functions of Matrices: Theory and Computation}, SIAM, Society for Industrial and Applied Mathematics (2008).

\bibitem{Horn1990}  Roger A. Horn, and Charles R. Johnson, 
\emph{Matrix analysis},
Cambridge University Press (1990).

\bibitem{Horn1991} Roger A. Horn, and Charles R. Johnson,
\emph{Topics in Matrix analysis}, $3^{rd}$ Edition.
Cambridge University Press (1991).

\bibitem{Mercer} Idris C. Mercer
\emph{Finding “nonobvious” nilpotent matrices}, (2005) \\
\texttt{http://www.idmercer.com/nilpotent.pdf}

\bibitem{Laub2005} Alan J. Laub,
\emph{Matrix analysis for scientists and engineers}, SIAM, Society for Industrial and Applied Mathematics (2005).

\bibitem{Meyer2000} Carl D. Meyer, 
\emph{Matrix analysis and applied linear algebra},
SIAM, Society for Industrial and Applied Mathematics (2000).

\bibitem{Strang2005} Gilbert Strang,
\emph{Linear Algebra and Its Applications}, SIAM, Society for Industrial and Applied Mathematics (2005).

\bibitem{Trefethen2000} Lloyd N. Trefethen, and David Bau,
\emph{Numerical linear algebra}, SIAM, Society for Industrial and Applied Mathematics (2000).

\bibitem{Eric} Eric W. Weisstein, "Characteristic Polynomial", from MathWorld--A Wolfram Web Resource.\\ \texttt{http://mathworld.wolfram.com/CharacteristicPolynomial.html}

\end{thebibliography}

\printindex

\end{document}