\chapter{Preface}

The Devil is in the details.

\section{Life is complex}
$\A{*}$ or $\A{T}$
  \begin{equation*}   %  =   =   =   =   =
  %\begin{split}
    \real{} \in \cmplx{}
    %\label{eq:}
  %\end{split}
  \end{equation*}

  \begin{table}[htbp]  %  T A B L E
    \caption{Matrix manipulations for $\A{*}$ and $\A{T}$.}
    \begin{center}
      \begin{tabular}{lll}
        %
        & \multicolumn{1}{c}{$\A{*}$} & \multicolumn{1}{c}{$\A{T}$} \\
        %
        Application & $\aicmn$ and $\A{}\in\real{\bymn}$ & only $\A{}\in\real{\bymn}$ \\\hline
        %
        Fortran     & $\texttt{conjg(transpose(A))}$  & $\texttt{transpose(A)}$ \\[5pt]
        %
        Mathematica & $\texttt{A}^{\texttt{H}}$ & $\texttt{A}^{\texttt{T}}$ \\
        %
                    & $\texttt{ConjugateTranspose[A]}$ & $\texttt{Transpose[A]}$ \\[5pt]
        %
        MATLAB      & $\texttt{A}$\textquotesingle  & $\texttt{A.}$\textquotesingle \\
        %
                    & $\texttt{ctranspose(A)}$ & $\texttt{transpose(A)}$ \\[5pt]
        %
        Python      & $\texttt{A.T.conj()}$ & $\texttt{A.T}$
        %
      \end{tabular}
    \end{center}
  %\label{tab:?}
  \end{table}%

\section{Readings}
There are many excellent books available examining many facets of the least squares problem. Fuller references are in the bibliography. \\

\noindent
\textbf{Carl Freidrich Gauss}\\
Theory of the Combination of Observations Least Subject to Errors

\section{Least Squares}  %    S    S    S    S    S    S    S    S    S
The titles are ranked by brevity. \\
 
\noindent
\textbf{Ilse C. F. Ipsen}\\
Numerical Matrix Analysis: Linear System and Least Squares (128 pp) \\

\noindent
\textbf{Charles L. Lawson}, and \textbf{Richard J. Hanson}\\
Solving Least Squares Problems (337 pp) \\

\noindent
\textbf{{\AA}ke Bj{\"o}rk}\\
Numerical Methods for Least Squares Problems (408 pp) \\

\section{Linear Algebra and Matrix Analysis}  %    S    S    S    S    S    S    S    S    S
The titles are ranked by brevity. \\

\noindent
\textbf{Alan J. Laub}\\
Matrix Analysis for Scientists and Engineers (157 pp) \\

\noindent
\textbf{Carl D. Meyer}\\
Matrix Analysis and Applied Linear Algebra (718 pp) \\

\section{Numerical Linear Algebra}  %    S    S    S    S    S    S    S    S    S
The titles are ranked by brevity. \\

\noindent
\textbf{Alan J. Laub}\\
Computational Matrix Analysis (154 pp) \\

\noindent
\textbf{Lloyd N. Trefethen}, and \textbf{David Bau, III} \\
Numerical Linear Algebra (361 pp) \\

\noindent
\textbf{E. Anderson}, \textbf{Z. Bai}, \textbf{C. Bischof}, \textbf{S. Blackford}, \textbf{J. Demmel}, \textbf{J. Dongarra}, \textbf{J Du Croz}, \textbf{A. Greenbaum}, \textbf{S. Hammarling}, \textbf{A. McKenney}, and \textbf{D. Sorenson} \\
LAPACK Users' Guide (407 pp)\\

\noindent
\textbf{G. W. Stewart}\\
Matrix Algorithms: \\
Volume I: Basic Decompositions (460 pp) \\
Volume II: Eigensystems (474 pp) \\

\section{Discussions on Least Squares}  %    S    S    S    S    S    S    S    S    S

Books with chapters dedicated to the topic. Sorted by author.\\

\noindent
\textbf{Gene H. Golub}, \textbf{Charles F. Van Loan} \\
Matrix Computations, ch. 5, 6 \\

\noindent
\textbf{Nicholas J. Higham} \\
Accuracy and Stability of Numerical Algorithms, ch. 20 \\

\noindent
\textbf{Cleve B. Moler} \\
Numerical Computing with MATLAB, ch. 5 \\

\noindent
\textbf{David S. Watkins} \\
Numerical Analysis: a mathematical introduction, ch. 5 \\

\noindent
\textbf{David S. Watkins} \\
Fundamentals of Matrix Computations, ch. 3 \\



\endinput