\chapter{Proofs of the singular value theorem}
Let's be candid: proofs have a checkered reputation in applied mathematics. Certainly, there are many docents who can wield a proof with skill and impart a sense of awe. But just as certainly, there are people who view proofs as detours to knowledge. Our assumption here is that the audience is application-oriented and some may require motivation to ponder proofs. For this reason, we delayed the proofs until presented the basic method and sound theoretic understanding. 

Libraries are full of books which start with proofs and we wanted to experiment with a variation on the theme. The Faustian bargain was this: to see how far we could get without proof. Hopefully, at this point your interest is piqued and you have an open mind towards the proofs. For those with inhibitions, one strategy is to work through the examples of the proofs.

The challenge is to show why people who want to compute things should care about proofs. The best summation we have heard boils to this simple statement:
\begin{quotation}
 A proof is worth an infinite number of exercises.
\end{quotation}
The point here is not to hop on the hamster-wheel of doing proofs to be able to prove more proves. It is to understand the applications in a more fundamental way which will build intuition. If you want to do a \svdl \  you can ask \emph{Mathematica} or Octave or use code from \emph{Numerical Recipes}. If you want to understand the limitations of the technique and how to apply it in your work, then you demand a deeper level of contact.

\input{chapters/"ch 12"/Laub}
\input{chapters/"ch 12"/improve}
\input{chapters/"ch 12"/constructive}

\endinput